---
title: "The basics first, you should learn"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The basics first, you should learn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  engine.path = '~/.virtualenvs/test/bin/python'
)
```

```{python}
import sys
print(sys.version)
```

```{r setup}
library(reservoir)
```

# Data

First let's generate awesome data

```{r}
n <- 10^2
awesome_data <- data.frame(t = 1:n,
                           x1 = rbinom(n = n, size = 1, prob = .5))
awesome_data$x2 <- sin(awesome_data$t/10)*2
awesome_data$y <- awesome_data$x2^2
```

```{r}
plot(awesome_data$t, awesome_data$y, type = "l")
```

# Echo state network setup

First we will define a reservoir network :

```{r}
reservoir <- reservoir::createNode(nodeType = "Reservoir", units = 100)
```

Then the readout set as a Ridge function :

```{r}
readout <- reservoir::createNode(nodeType = "Ridge", ridge = 0.01)
```

And we connect them :

```{r}
model <- reservoir::link(reservoir, readout)
```

# Train the model and predict

Let's try to predict 7 time points ahead then :

```{r}
# train set
yTrain <- awesome_data$y[8:50]
xTrain <- awesome_data[1:43, c("x1", "x2")]
# test set
xTest <- awesome_data[44:(n-7), c("x1", "x2")]
yTest <- awesome_data$y[51:n]
```

Before to fit the model we have to convert them to array :

```{r}
lsdf <- lapply(list(yTrain = yTrain,
            xTrain = xTrain,
            xTest = xTest,
            yTest = yTest),
       function(x) as.array(as.matrix(x)))
```

And now let's fit the model

```{r}
fit <- reservoir::fit(node = model, X = lsdf$xTrain, Y = lsdf$yTrain)
```

And predict

```{r}
vecPredictions = reservoir::predict_seq(node = fit, X = lsdf$xTest)
```

We can compare the prediction and the truth :

```{r}
plot(awesome_data$t[51:n], yTest, type = "l")
lines(awesome_data$t[51:n], vecPredictions, col = "red")
```

